Classification structurée pour l'apprentissage par renforcement

### 2012-09-14 klein2012structured_RIA/scirl_ria.tex
Réglé un problème particulièrement têtu avec les figures

\firstpagenumber{9}

@@

\begin{document}

\maketitle

\newpage
\cite{Ng:2000}

\section{Introduction}

\acknowledgements{...}

\bibliography{scirl_ria.bib}

>>>>

\firstpagenumber{9}
\let\chapter\section
\usepackage[ruled,algo2e]{algorithm2e}
\usepackage{mathtools}
\usepackage{subfigure}
\mathtoolsset{showonlyrefs=true}

\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\argmin}{\operatorname*{argmin}}
\newcommand{\diag}{\operatorname*{diag}}

\newcommand{\p}{\mathcal{P}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\s}{\mathcal{S}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\lc}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\Mu}{\boldsymbol{\mu}}
%\newcommand{\Xi}{\boldsymbol{\xi}}

%\newtheorem{theorem}{Theorem}

@@

\begin{document}
\newtheorem{prop}[theorem]{Proposition}
\cite{Ng:2000}
\maketitle

\newpage

\section{Background and Notations}
\label{sec:background}

\subsection{(Inverse) Reinforcement Learning}
\label{subsec:background:irl}

A Markov Decision process (MDP)~\cite{Puterman:1994} is a tuple
$\{\s,\A,\p,\R,\gamma\}$ where $\s$ is the finite state
space\footnote{This work can be extended to compact state spaces, up
to some technical aspects.}, $\A$ the finite actions space, $\p =
\{P_a = (p(s'|s,a))_{1\leq s,s'\leq |\s|}, a\in\A\}$ the set of
Markovian transition probabilities, $\R\in\mathbb{R}^\s$ the
state-dependent reward function and $\gamma$ the discount factor. A
deterministic policy $\pi\in\s^\A$ defines the behavior of an agent.
The quality of this control is quantified by the value function
$v_\R^\pi\in\mathbb{R}^\s$, associating to each state the cumulative
discounted reward for starting in this state and following the
policy $\pi$ afterwards: $v_\R^\pi(s) = \E[\sum_{t\geq 0} \gamma^t
\R(S_t)|S_0=s,\pi]$. An optimal policy $\pi_\R^*$ (according to the
reward function $\R$) is a policy of associated value function
$v^*_\R$ satisfying $v_\R^* \geq v_\R^\pi$, for any policy $\pi$ and
componentwise.

Let $P_\pi$ be the stochastic matrix $P_\pi =
(p(s'|s,\pi(s)))_{1\leq s,s'\leq |\s|}$. With a slight abuse of
notation, we may write $a$ the policy which associates the action
$a$ to each state $s$. The Bellman evaluation (resp. optimality)
operators $T^\pi_\R\text{ (resp. $T^*_\R$)}:\mathbb{R}^\s
\rightarrow \mathbb{R}^\s$ are defined as $T^\pi_\R v = \R + \gamma
P_\pi v$ and $T_\R^*v = \max_\pi T_\R^\pi v$.
%\begin{equation}
%  T^\pi_\R v = \R + \gamma P_\pi v \text{ and } T_\R^*v = \max_\pi T_\R^\pi v.
%\end{equation}
These operators are contractions and $v_\R^\pi$ and $v^*_\R$ are
their respective fixed-points: $v_\R^\pi = T^\pi_\R v_\R^\pi$ and
$v^*_\R = T^*_\R v^*_\R$. The action-value function
$Q^\pi\in\mathbb{R}^{\s\times \A}$ adds a degree of freedom on the
choice of the first action, it is formally defined as $Q_\R^\pi(s,a)
= [T^a_\R v^\pi_\R](s)$. We also write $\rho_\pi$ the stationary
distribution of the policy $\pi$ (satisfying $\rho_\pi^\top P_\pi =
\rho_\pi^\top$).

Reinforcement learning and approximate dynamic programming aim at
estimating the optimal control policy $\pi^*_\R$ when the model
(transition probabilities and the reward function) is unknown (but
observed through interactions with the system to be controlled) and
when the state space is too large to allow exact representations of
the objects of interest (as value functions or
policies)~\cite{Bertsekas:1996,Sutton:1998,szepesvari2010c}.
%
We refer to this as the direct problem.
%
On the contrary, (approximate) inverse reinforcement
learning~\cite{Ng:2000} aim at estimating a reward function for
which an observed policy is (nearly) optimal.
%
Let us call this policy the expert policy, denoted $\pi_E$. We may
assume that it optimizes some unknown reward function $\R_E$. The
aim of IRL is to compute some reward $\hat{\R}$ such that the expert
policy is (close to be) optimal, that is such that $v^*_{\hat{\R}}
\approx v^{\pi_E}_{\hat{\R}}$.
%
We refer to this as the inverse problem.

Similarly to the direct problem, the state space may be too large
for the reward function to admit a practical exact representation.
Therefore, we restrict our search of a good reward among linearly
parameterized functions. Let $\phi(s) = (\phi_1(s)  \dots
\phi_p(s))^\top$
%\begin{pmatrix}
%  \phi_1(s) & \dots & \phi_p(s)
%\end{pmatrix}^\top$
be a feature vector composed of $p$ basis function
$\phi_i\in\mathbb{R}^\s$, we define the  parameterized reward
functions as $\R_\theta(s) = \theta^\top \phi(s) = \sum_{i=1}^p
\theta_i \phi_i(s)$.
%\begin{equation}
%  \R_\theta(s) = \theta^\top \phi(s) = \sum_{i=1}^p \theta_i \phi_i(s).
%\end{equation}
Searching a good reward thus reduces to searching a good parameter
vector $\theta \in\mathbb{R}^p$. Notice that we will use
interchangeably $\R_\theta$ and $\theta$ as subscripts
(\textit{e.g.}, $v_\theta^\pi$ for $v_{\R_\theta}^\pi$).
Parameterizing the reward this way implies a related
parameterization for the action-value function:
\begin{equation}
  Q^\pi_\theta(s,a) = \theta^\top \mu^\pi(s,a) \text{ with }
  \mu^\pi(s,a) = \E[\sum_{t\geq 0} \gamma^t
  \phi(S_t)|S_0=s,A_0=a,\pi].
  \label{eq:def:mu}
\end{equation}
Therefore, the action-value function shares the parameter vector of
the reward function, with an associated feature vector $\mu^\pi$
called the \emph{feature expectation}. This notion will be of
primary importance for the contribution of this paper. Notice that
each component $\mu_i^\pi$ of this feature vector is actually the
action-value function of the policy $\pi$ assuming the reward is
$\phi_i$: $\mu_i^\pi(s,a) = Q^\pi_{\phi_i}(s,a)$. Therefore, any
algorithm designed for estimating an action-value function may be
used to estimate the feature expectation, such as Monte-Carlo
rollouts or temporal difference learning~\cite{Klein:2011}.

\subsection{Classification with Linearly Parameterized Score
Functions} \label{subsec:background:classif}

Let $\X$ be a compact or a finite set (of inputs to be classified)
and let $\Y$ be a finite set (of labels). Assume that inputs $x\in
\X$  are drawn according to some unknown distribution $\prob(x)$ and
that there exists some oracle which associates to each of these
inputs a label $y\in Y$ drawn according to the unknown conditional
distribution $\prob(y|x)$. Generally speaking, the goal of
multi-class classification is, given a training set
$\{(x_i,y_i)_{1\leq i \leq N}\}$ drawn according to $\prob(x,y)$, to
produce a decision rule $g\in\Y^\X$ which aims at minimizing the
classification error $\E[\chi_{\{g(x)\neq y\}}] = \prob(g(x)\neq
y)$, where $\chi$ denotes the indicator function.

Here, we consider a more restrictive set of classification
algorithms. We assume that the decision rule associates to an input
the argument which maximizes a related score function, this score
function being linearly parameterized and the associated parameters
being learnt by the algorithm. More formally, let $\psi(s,a) =
(\psi_1(x,y)  \dots  \psi_d(x,y))^\top\in \mathbb{R}^d$
%\begin{pmatrix}   \psi_1(x,y) & \dots & \psi_d(x,y)
%\end{pmatrix}^\top \in \mathbb{R}^d$
be a feature vector whose components are $d$ basis functions
$\psi_i\in\mathbb{R}^{\X\times\Y}$. The linearly parameterized score
function $s_w\in\mathbb{R}^{\X\times \Y}$ of parameter vector
$w\in\mathbb{R}^d$ is defined as $s_w(x,y) = w^\top \psi(x,y)$. The
associated decision rule $g_w\in{\Y^\X}$ is defined as $g_w(x) \in
\argmax_{y\in\Y}s_w(x,y)$. Using a training set $\{(x_i,y_i)_{1\leq
i\leq N}\}$, a linearly parameterized score function-based
multi-class classification (MC$^2$ for short) algorithm computes a
parameter vector $\theta_c$. The quality of the solution is
quantified by the classification error $\epsilon_c =
\prob(g_{\theta_c}(x)\neq y)$.

We do not consider a specific MC$^2$ algorithm, as long as it
classifies inputs by maximizing the argument of a linearly
parameterized score function. For example, one may choose a
multi-class support vector machine~\cite{Guermeur:2007} (taking the
kernel induced by the feature vector) or a structured large margin
approach~\cite{Taskar:2005}. Other choices may be possible, one can
choose its preferred algorithm.

\section{Structured Classification for Inverse Reinforcement
Learning} \label{sec:scirl}

\subsection{General Algorithm}
\label{subsec:scirl:algo}

Consider the classification framework of
Sec.~\ref{subsec:background:classif}. The input $x$ may be seen as a
state and the label $y$ as an action. Then, the decision rule
$g_w(x)$ can be interpreted as a policy which is greedy according to
the score function $w^\top \psi(x,y)$, which may itself be seen as
an action-value function. Making the parallel with
Eq.~\eqref{eq:def:mu}, if $\psi(x,y)$ is the feature expectation of
some policy $\pi$ which produces labels of the training set, and if
the classification error is small, then $w$ will be the parameter
vector of a reward function for which we may hope the policy $\pi$
to be near optimal. Based on these remarks, we're ready to present
the proposed Structured Classification-based IRL (SCIRL) algorithm.

Let $\pi_E$ be the expert policy from which we would like to recover
a reward function. Assume that we have a training set $\D =
\{(s_i,a_i=\pi_E(s_i))_{1\leq i\leq N}\}$ where states are sampled
according to the expert stationary distribution\footnote{For
example, if the Markov chain induced by the expert policy is
fast-mixing, sampling a trajectory will quickly lead to sample
states according to this distribution.} $\rho_E = \rho_{\pi_E}$.
Assume also that we have an estimate $\hat{\mu}^{\pi_E}$ of the
expert feature expectation $\mu^{\pi_E}$ defined in
Eq.~\eqref{eq:def:mu}. How to practically estimate this quantity is
postponed to Sec.~\ref{subsec:practicalApproach:muE}; however,
recall that estimating $\mu^{\pi_E}$ is simply a \emph{policy
evaluation} problem (estimating the action-value function of a given
policy), as noted in Sec.~\ref{subsec:background:irl}. Assume also
that an MC$^2$ algorithm has been chosen. The proposed algorithm
simply consists in choosing $\theta^\top\hat{\mu}^{\pi_E}(s,a)$ as
the linearly parameterized score function, training the classifier
on $\D$ which produces a parameter vector $\theta_c$, and outputting
the reward function $\R_{\theta_c}(s) = \theta_c^\top \phi(s)$.

\begin{algorithm2e}%[tbh]
    %\small
  \SetAlgoVlined
  \caption{SCIRL algorithm}
  \label{algo:scirl}
  %
  \BlankLine
  \emph{\textbf{Given}} a training set $\D = \{(s_i,a_i=\pi_E(s_i))_{1\leq i\leq N}\}$,
  an estimate $\hat{\mu}^{\pi_E}$ of the expert feature expectation $\mu^{\pi_E}$ and an MC$^2$ algorithm\;
  %
  \BlankLine
  \emph{\textbf{Compute}} the parameter vector $\theta_c$ using the
  MC$^2$ algorithm
  fed with the training set $\D$ and considering the parameterized score function
  $\theta^\top\hat{\mu}^{\pi_E}(s,a)$\;
  %
  \BlankLine
  \emph{\textbf{Output}} the reward function $\R_{\theta_c}(s) = \theta_c^\top\phi(s)$ \;
\end{algorithm2e}

The proposed approach is summarized in Alg.~\ref{algo:scirl}. We
call this Structured Classification-based IRL because using the
(estimated) expert feature expectation as the feature vector for the
classifier somehow implies taking into account the MDP structure
into the classification problem and allows outputting a reward
vector. Notice that contrary to most of existing IRL
%and apprenticeship learning
algorithms, SCIRL does not require solving the direct problem. If it
possibly requires estimating the expert feature expectation, it is
just a policy evaluation problem, less difficult than the policy
optimization issue involved by the direct problem. This is further
discussed in Sec.~\ref{sec:relatedWorks}.
%
%Now, we formally show that the proposed approach makes sens, as long
%as $\mu^{\pi_E}$ is well estimated and as long as the classification
%error is small.

\subsection{Analysis}
\label{subsec:scirl:analysis}

In this section, we show that the expert policy $\pi_E$ is close to
be optimal according to the reward function $\R_{\theta_c}$, more
precisely that
$\E_{s\sim\rho_E}[v^*_{\theta_c}(s)-v^{\pi_E}_{\theta_c}(s)]$ is
small. Before stating our main result, we need to introduce some
notations and to define some objects.

We will use the first order discounted future state distribution
concentration coefficient $C_f$~\cite{Munos:2007}:
\begin{equation}
  C_f = (1-\gamma)\sum_{t\geq 0} \gamma^t c(t) \text{ with } c(t) =
  \max_{\pi_1,\dots,\pi_t,s\in\s}\frac{(\rho_E^\top P_{\pi_1}\dots
  P_{\pi_t})(s)}{\rho_E(s)}.
\end{equation}
We note $\pi_c$ the decision rule of the classifier: $\pi_c(s) \in
\argmax_{a\in \A} \theta_c^\top\hat{\mu}^{\pi_E}(s,a)$. The
classification error is therefore $\epsilon_c =
\E_{s\sim\rho_E}[\chi_{\{\pi_c(s)\neq\pi_E(s)\}}] \in [0,1]$. We
write $\hat{Q}^{\pi_E}_{\theta_c} = \theta_c^\top \hat{\mu}^{\pi_E}$
the score function computed from the training set $\D$ (which can be
interpreted as an approximate action-value function). Let also
$\epsilon_{\mu} = \hat{\mu}^{\pi_E} - \mu^{\pi_E}:\s\times\A
\rightarrow  \mathbb{R}^p$ be the feature expectation error.
Consequently, we define the action-value function error as
$\epsilon_Q = \hat{Q}^{\pi_E}_{\theta_c} - Q^{\pi_E}_{\theta_c} =
\theta_c^\top(\hat{\mu}^{\pi_E} - \mu^{\pi_E}) = \theta_c^\top
\epsilon_\mu:\s\times\A\rightarrow\mathbb{R}$. We finally define the
mean delta-max action-value function error as $\bar{\epsilon}_Q =
\E_{s\sim\rho_E}[\max_{a\in\A}\epsilon_Q(s,a) -
\min_{a\in\A}\epsilon_Q(s,a)]\geq 0$.

\begin{theorem}
  \label{th}
  Let $\R_{\theta_c}$ be the reward function outputted by
  Alg.~\ref{algo:scirl}. Let also the quantities $C_f$, $\epsilon_c$
  and $\bar{\epsilon}_Q$ be defined as above. We have
  %
%  the concentration coefficient $C_f$,
%  the classification error $\epsilon_c$ and the mean-max
%  action-value function error $\bar{\epsilon}_Q$ be defined as
%  above. Then, we have
  \begin{equation}
    0\leq
    \E_{s\sim\rho_E}[v^*_{\R_{\theta_c}}-v^{\pi_E}_{\R_{\theta_c}}]
    \leq \frac{C_f}{1-\gamma}\left(\bar{\epsilon}_Q +
    \epsilon_c\frac{2\gamma\|\R_{\theta_c}\|_\infty}{1-\gamma}
%    \epsilon_c\frac{2\gamma}{1-\gamma}\min(\|\R_{\theta_c}\|_\infty,
%    C_p \E_{s\sim\rho_E}[\R_{\theta_c}(s)])
    \right).
  \end{equation}
\end{theorem}
\begin{proof}
  As the proof only relies on the reward $\R_{\theta_c}$, we
  omit the related subscripts to keep the notations simple
  (\textit{e.g.}, $v^\pi$ for
  $v^\pi_{\theta_c}=v^\pi_{\R_{\theta_c}}$ or $\R$ for $\R_{\theta_c}$). First, we link
  the error $\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)]$ to the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s)-v^{\pi_E}(s)]$.
  %
  % A second step consists in bounding this residual.
  %
  Componentwise, we have that:
  \begin{align}
    v^* - v^{\pi_E} &= T^* v^*  - T^{\pi^*}v^{\pi_E} +
    T^{\pi^*}v^{\pi_E} - T^* v^{\pi_E} + T^* v^{\pi_E} - v^{\pi_E}
    \\
    &\stackrel{(a)}{\leq} \gamma P_{\pi^*}(v^*-v^{\pi_E}) + T^*
    v^{\pi_E} - v^{\pi_E}
    %
    \stackrel{(b)}{\leq} (I-\gamma
    P_{\pi^*})^{-1} (T^* v^{\pi_E} - v^{\pi_E}).
  \end{align}
  Inequality $(a)$ holds because $T^{\pi^*} v^{\pi_E}\leq T^*
  v^{\pi_E}$ and inequality $(b)$ holds thanks
  to~\cite[Lemma~4.2]{Munos:2007}. Moreover, $v^*$ being optimal we
  have that $v^*-v^{\pi_E}\geq 0$ and $T^*$ being the Bellman
  optimality operator, we have $T^* v^{\pi_E}\geq
  T^{\pi_E}v^{\pi_E}=v^{\pi_E}$. Additionally, remark that
  $(I-\gamma P_{\pi^*})^{-1} = \sum_{t\geq 0}\gamma^t P_{\pi^*}^t$.
  Therefore, using the definition of the concentration coefficient
  $C_f$, we have that:
  \begin{equation}
    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq \frac{C_f}{1-\gamma}
    \E_{s\sim\rho_E}\left[[T^*v^{\pi_E}](s) - v^{\pi_E}(s)\right].
    \label{eq:proof:residual}
  \end{equation}
  This results actually follows closely the one
  of~\cite[Theorem~4.2]{Munos:2007}. There remains to bound the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s) -
  v^{\pi_E}(s)]$. Considering the following decomposition,
  \begin{equation}
    T^* v^{\pi_E} - v^{\pi_E} = T^* v^{\pi_E} - T^{\pi_c}v^{\pi_E}
    + T^{\pi_c}v^{\pi_E}- v^{\pi_E},
    \label{eq:proof:decomposition}
  \end{equation}
  we will bound $\E_{s\sim\rho_E}[[T^* v^{\pi_E}](s) - [T^{\pi_c}v^{\pi_E}](s)]$
  and $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]$.

  The policy $\pi_c$ (the decision rule of the classifier) is greedy
  with respect to
  $\hat{Q}^{\pi_E}=\theta_c^\top\hat{\mu}^{\pi_E}$. Therefore, for any state-action couple
  $(s,a)\in\s\times \A$ we have:
  \begin{equation}
    \hat{Q}^{\pi_E}(s,\pi_c(s))\geq
    \hat{Q}^{\pi_E}(s,a)
    \Leftrightarrow
    Q^{\pi_E}(s,a) \leq Q^{\pi_E}(s,\pi_c(s)) +
    \epsilon_Q(s,\pi_c(s)) - \epsilon_Q(s,a).
  \end{equation}
  By definition, $Q^{\pi_E}(s,a) = [T^a v^{\pi_E}](s)$ and
  $Q^{\pi_E}(s,\pi_c(s)) = [T^{\pi_c} v^{\pi_E}](s)$. Therefore, for $s\in\s$:
  \begin{align}
    \forall a\in A,\; [T^a v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \epsilon_Q(s,\pi_c(s))-\epsilon_Q(s,a)
    \\
    \Rightarrow [T^* v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \max_{a\in \A}\epsilon_Q(s,a)-\min_{a\in
    \A}\epsilon_Q(s,a).
  \end{align}
  Taking the expectation according to $\rho_E$ and noticing that
  $T^* v^{\pi_E}\geq v^{\pi_E}$, we bound the first
  term:
  \begin{equation}
    0 \leq \E_{s\sim\rho_E}\left[ [T^* v^{\pi_E}](s) - [T^{\pi_c}
    v^{\pi_E}](s)\right] \leq \bar{\epsilon}_Q.
    \label{eq:proof:b1}
  \end{equation}
  There finally remains to bound the term $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) -
  v^{\pi_E}(s)]$.

  Let us write $M\in\mathbb{R}^{|\s|\times |\s|}$ the diagonal matrix
  defined as $M = \diag (\chi_{\{\pi_c(s)\neq\pi_E(s)\}})$. Using
  this, the Bellman operator $T^{\pi_c}$ may be written as, for any $v\in\mathbb{R}^\s$:
  \begin{equation}
    T^{\pi_c}v = \R + \gamma M P_{\pi_c} v + \gamma (I-M)P_{\pi_E} v
    = \R + \gamma P_{\pi_E} v + \gamma M (P_{\pi_c}-P_{\pi_E})v.
  \end{equation}
  Applying this operator to $v^{\pi_E}$ and recalling that $\R +
  \gamma P_{\pi_E} v^{\pi_E} = T^{\pi_E} v^{\pi_E} = v^{\pi_E}$, we
  get:
  \begin{equation}
    T^{\pi_c}v^{\pi_E} - v^{\pi_E} = \gamma M
    (P_{\pi_c}-P_{\pi_E})v^{\pi_E}
    \Rightarrow |\rho_E^\top (T^{\pi_c}v^{\pi_E} - v^{\pi_E})| = \gamma
    |\rho_E^\top M (P_{\pi_c}-P_{\pi_E})v^{\pi_E}|.
  \end{equation}
  One can easily see that $\|(P_{\pi_c}-P_{\pi_E})v^{\pi_E}\|_\infty
  \leq \frac{2}{1-\gamma}\|\R\|_\infty$, which allows bounding the
  last term:
  \begin{equation}
    |\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]| \leq
    \epsilon_c \frac{2\gamma}{1-\gamma} \|\R\|_\infty.
    \label{eq:proof:b2}
  \end{equation}
  Injecting bounds of Eqs.~\eqref{eq:proof:b1}
  and~\eqref{eq:proof:b2} into Eq.~\eqref{eq:proof:residual} gives the
  stated result.%, which concludes the proof.
%  Injecting bounds of Eqs.~\eqref{eq:proof:b1}
%  and~\eqref{eq:proof:b2} into Eq.~\eqref{eq:proof:residual} using the
%  decomposition of Eq.~\eqref{eq:proof:decomposition} gives the
%  stated result:
%  \begin{equation}
%    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq
%    \frac{C_f}{1-\gamma} \left(\bar{\epsilon}_Q + \epsilon_c \frac{2\gamma}{1-\gamma} \|\R\|_\infty
%    \right).
%  \end{equation}
%  This concludes the proof.
\end{proof}

This result shows that if the expert feature expectation is well
estimated (in the sense that the estimation error $\epsilon_\mu$ is
small for states sampled according to the expert stationary policy
and for all actions) and if the classification error $\epsilon_c$ is
small, then the proposed generic algorithm outputs a reward function
$\R_{\theta_c}$ for which the expert policy will be near optimal. A
direct corollary of Th.~\ref{th} is that given the true expert
feature expectation $\mu^{\pi_E}$ and a perfect classifier
($\epsilon_c=0$), $\pi_E$ is the unique optimal policy for
$\R_{\theta_c}$.

One may argue that this bounds trivially holds for the null reward
function (a reward often exhibited to show that IRL is an ill-posed
problem), obtained if $\theta_c=0$. However, recall that the
parameter vector $\theta_c$ is computed by the classifier. With
$\theta_c=0$, the decision rule would be a random policy and we
would have $\epsilon_c = \frac{|\A|-1}{|\A|}$, the worst possible
classification error. This case is really unlikely. Therefore, we
advocate that the proposed approach somehow allows disambiguating
the IRL problem (at least, it does not output trivial reward
functions such as the null vector).
%
Also, this bound is scale-invariant: one could impose
$\|\theta_c\|=1$ or normalize (action-) value functions by
$\|\R_{\theta_c}\|_\infty^{-1}$.

One should notice that there is a hidden dependency of the
classification error $\epsilon_c$ to the estimated expert feature
expectation $\hat{\mu}^{\pi_E}$. Indeed, the minimum classification
error depends on the hypothesis space spanned by the chosen score
function basis functions for the MC$^2$ algorithm (here
$\hat{\mu}^{\pi_E}$). Nevertheless, provided a good representation
for the reward function (that is a good choice of basis functions
$\phi_i$) and a small estimation error, this should not be a
practical problem.

Finally, if our bound relies on the generalization errors
$\epsilon_c$ and $\bar{\epsilon}_Q$, the classifier will only use
$(\hat{\mu}^{\pi_E}(s_i,a))_{1\leq i\leq N,a\in\A}$ in the training
phase, where $s_i$ are the states from the set $\D$. It outputs
$\theta_c$, seen as a reward function, thus the estimated feature
expectation $\hat{\mu}^{\pi_E}$ is no longer required. Therefore,
practically it should be sufficient to estimate well
$\hat{\mu}^{\pi_E}$ on state-action couples $(s_i,a)_{1\leq i\leq
N,a\in\A}$, which allows envisioning Monte-Carlo rollouts for
example.

\section{A Practical Approach}
\label{sec:practicalApproach}

\subsection{Estimating the Expert Feature Expectation}
\label{subsec:practicalApproach:muE}

SCIRL relies on an estimate $\hat{\mu}^{\pi_E}$ of the expert
feature expectation. Basically, this is a policy evaluation problem.
An already made key observation is that each component of
$\mu^{\pi_E}$ is the action-value function of $\pi_E$ for a reward
function $\phi_i$: $\mu_i^{\pi_E}(s,a) = Q^{\pi_E}_{\phi_i}(s,a) =
[T^a_{\phi_i} v^{\pi_E}_{\phi_i}](s)$. We briefly review its exact
computation and possible estimation approaches, and consider
possible heuristics.
%possibly considering some heuristics.

If the model is known, the feature expectation can be computed
explicitly. Let $\Phi\in\mathbb{R}^{|\s|\times p}$ be the feature
matrix whose rows contain the feature vectors $\phi(s)^\top$ for all
$s\in\s$.
%
%(equivalently, whose columns contain the basis functions $\phi_i$
%for $1\leq i\leq p$).
%
For a fixed $a\in A$, let $\Mu^{\pi_E}_a \in\mathbb{R}^{|\s|\times
p}$ be the feature expectation matrix whose rows are the expert
feature vectors, that is $(\mu^{\pi_E}(s,a))^\top$ for any $s\in\s$.
With these notations, we have $\Mu_a^{\pi_E} = \Phi + \gamma
P_a(I-\gamma P_{\pi_E})^{-1} \Phi$.
%\begin{equation}
%  \forall a\in\A,\; \Mu_a^{\pi_E} = \Phi + \gamma P_a(I-\gamma
%  P_{\pi_E})^{-1} \Phi.
%\end{equation}
%Therefore, knowing the model (the dynamic), the expert feature
%expectation can be easily computed.
Moreover, the related computational cost is the same order of
magnitude as evaluating a single policy (as the costly part,
computing $(I-\gamma P_{\pi_E})^{-1}$, is shared by all components).

If the model is unknown, any temporal difference learning algorithm
can be used to estimate the expert feature
expectation~\cite{Klein:2011}, as LSTD (Least-Squares Temporal
Differences)~\cite{Bradtke:1996}. Let $\psi:\s\times \A \rightarrow
\mathbb{R}^d$ be a feature vector composed of $d$ basis functions
$\psi_i \in\mathbb{R}^{\s\times \A}$. Each component $\mu_i^{\pi_E}$
of the expert feature expectation is parameterized by a vector
$\xi_i\in\mathbb{R}^d$: $\mu_i^{\pi_E}(s,a)\approx \xi_i^\top
\psi(s,a)$. Assume that we have a training set
$\{(s_i,a_i,s'_i,a'_i=\pi_E(s'_i))_{1\leq i \leq M}\}$ with actions
$a_i$ not necessarily sampled according to policy $\pi_E$
(\textit{e.g.}, this may be obtained by sampling trajectories
according to an expert-based $\epsilon$-greedy policy), the aim
being to have a better variability of tuples (non-expert actions
should be tried). Let $\tilde{\Psi}\in \mathbb{R}^{M\times d}$
(resp. $\tilde{\Psi}'$) be the feature matrix whose rows are the
feature vectors $\psi(s_i,a_i)^\top$ (resp. $\psi(s'_i,a'_i)^\top$).
Let also $\tilde{\Phi}\in \mathbb{R}^{M\times p}$ be the feature
matrix whose rows are the reward's feature vectors $\phi(s_i)^\top$.
Finally, let $\Xi =
\begin{bmatrix}   \xi_1 & \dots & \xi_p
\end{bmatrix}\in\mathbb{R}^{d\times p}$ be the matrix of all
parameter vectors. Applying LSTD to each component of the feature
expectation gives the LSTD-$\mu$ algorithm~\cite{Klein:2011}: $\Xi =
(\tilde{\Psi}^\top(\tilde{\Psi} - \gamma
\tilde{\Psi}'))^{-1}\tilde{\Psi}^\top \tilde{\Phi}$ and
$\hat{\mu}^{\pi_E}(s,a) = \Xi^\top \psi(s,a)$.
%\begin{equation}
%  \Xi = \left(\tilde{\Psi}^\top(\tilde{\Psi} - \gamma
%  \tilde{\Psi}')\right)^{-1}\tilde{\Psi}^\top \tilde{\Phi}
%  \text{ and } \hat{\mu}^{\pi_E}(s,a) = \Xi^\top \psi(s,a).
%\end{equation}
As for the exact case, the costly part (computing the inverse
matrix) is shared by all feature expectation components, the
computational cost is reasonable (same order as LSTD).

%
Provided a simulator and the ability to sample according to the
expert policy, the expert feature expectation may also be estimated
using Monte-Carlo rollouts for a given state-action pair (as noted
in Sec.~\ref{subsec:scirl:analysis}, $\hat{\mu}^{\pi_E}$ need only
be known on $(s_i,a)_{1\leq i\leq N,a\in\A}$). Assuming that $K$
trajectories are sampled for each required state-action pair, this
method would require $KN|\A|$ rollouts.

In order to have a small error $\bar{\epsilon}_Q$, one may learn
using transitions whose starting state is sampled according to
$\rho_E$ and whose actions are uniformly distributed. However, it
may happen that only transitions of the expert are available: $\T =
\{(s_i,a_i=\pi_E(s_i),s'_i)_{1\leq i \leq N}\}$. If the state-action
couples $(s_i,a_i)$ may be used to feed the classifier, the
transitions $(s_i,a_i,s'_i)$ are not enough to provide an accurate
estimate of the feature expectation. In this case, we can still
expect an accurate estimate of $\mu^{\pi_E}(s,\pi_E(s))$, but there
is little hope for $\mu^{\pi_E}(s,a\neq\pi_E(s))$. However, one can
still rely on some heuristic; this does not fit the analysis of
Sec.~\ref{subsec:scirl:analysis}, but it can still provide good
experimental results, as illustrated in Sec.~\ref{sec:experiments}.

We propose such a heuristic. Assume that only data $\T$ is available
and that we use it to provide an (accurate) estimate
$\hat{\mu}^{\pi_E}(s,\pi_E(s))$ (this basically means estimating a
value function instead of an action-value function as described
above). We may adopt an optimistic point of view by assuming that
applying a non-expert action just delays the effect of the expert
action. More formally, we associate to each state $s$ a virtual
state $s_\text{v}$ for which $p(.|s_\text{v},a)=p(.|s,\pi_E(s))$ for
any action $a$ and for which the reward feature expectation is the
null vector, $\phi(s_v) = 0$. In this case, we have
$\mu^{\pi_E}(s,a\neq\pi_E(s)) = \gamma \mu^{\pi_E}(s,\pi_E(s))$.
Applying this idea to the available estimate (recalling that the
classifiers only requires evaluating $\hat{\mu}^{\pi_E}$ on
$(s_i,a)_{1\leq i\leq N,a\in \A}$) provides the proposed heuristic:
for $1\leq i\leq N$, $\hat{\mu}^{\pi_E}(s_i,a\neq a_i) = \gamma
\hat{\mu}^{\pi_E}(s_i,a_i)$.
%\begin{equation}
%  \forall 1\leq i \leq N,\;
%  \hat{\mu}^{\pi_E}(s_i,a\neq a_i) = \gamma
%  \hat{\mu}^{\pi_E}(s_i,a_i).
%  \label{eq:heuristic}
%\end{equation}

We may even push this idea further, to get the simpler estimate of
the expert feature expectation (but with the weakest guarantees).
Assume that the set $\T$ consists of one long trajectory, that is
$s'_i = s_{i+1}$ (thus $\T =
\{s_1,a_1,s_2,\dots,s_{N-1},a_{N-1},s_N,a_N\}$). We may estimate
$\mu^{\pi_E}(s_i,a_i)$ using the single rollout available in the
training set and use the proposed heuristic for other actions:
\begin{equation}
  \forall 1\leq i \leq N,\; \hat{\mu}^{\pi_E}(s_i,a_i) =
  \sum_{j=i}^N \gamma^{j-i}\phi(s_j) \text{ and }
  \hat{\mu}^{\pi_E}(s_i,a\neq a_i) = \gamma
  \hat{\mu}^{\pi_E}(s_i,a_i).
  \label{eq:mc_plus_heuristic}
\end{equation}

To sum up, the expert feature expectation may be seen as a vector of
action-value functions (for the same policy $\pi_E$ and different
reward functions $\phi_i$). Consequently, any action-value function
evaluation algorithm may be used to estimate $\mu^\pi(s,a)$.
Depending on the available data, one may have to rely on some
heuristic to assess the feature expectation for a unexperienced
(non-expert) action. Also, this expert feature expectation estimate
is only required for training the classifier, so it is sufficient to
estimate on state-action couples $(s_i,a)_{1\leq i \leq N,a\in \A}$.
In any case, estimating $\mu^{\pi_E}$ is not harder than estimating
the action-value function of a given policy in the on-policy case,
which is much easier than computing an optimal policy for an
arbitrary reward function (as required by most of existing IRL
algorithms, see Sec.~\ref{sec:relatedWorks}).

\subsection{An Instantiation}
\label{subsec:practicalApproach:instantiation}
